---
title: "Assignment 6"
author: "Grant Jackson"
date: "2024-11-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(fixest)
library(DRDID)
```

```{r load data}
setwd("~/Applied Microeconometrics/Data")

tva <- read_csv('tva.csv')
```

## Question 1: Different DiD Estimation Methods

```{r creating short data}
tva_short <- tva |> 
  filter(year == 1940 | year == 1960)

tva_short
```

### Method 1

```{r did-manual}
ybar_treat_post <- tva |>
  filter(year == 1960 & tva == 1) |> 
  pull(ln_manufacturing) |> mean()

ybar_treat_pre <- tva |>
  filter(year == 1940 & tva == 1) |>
  pull(ln_manufacturing) |> mean()

ybar_control_post <- tva |>
  filter(year == 1960 & tva == 0) |> 
  pull(ln_manufacturing) |> mean()

ybar_control_pre <- tva |>
  filter(year == 1940 & tva == 0) |>
  pull(ln_manufacturing) |> mean()

did_manual <- (ybar_treat_post - ybar_treat_pre) - (ybar_control_post - ybar_control_pre)

print(paste("Manual DiD Estimate:", did_manual))
```

### Method 2

```{r did-regression}
reg_did <- feols(
  ln_manufacturing ~
    i(tva, year == 1960, ref = 0) | county_code + year,
  data = tva_short, vcov = "hc1"
)
    

print(reg_did)
```

### Method 3

```{r did-firstdiff}
first_diff <- tva_short |>
  mutate(
    .by = county_code,
    delta_ln_manufacturing =
      ln_manufacturing[year == 1960] -
      ln_manufacturing[year == 1940]
  ) |>
  filter(year == 1960)

did_fd <- feols(
  delta_ln_manufacturing ~ i(tva, ref = 0),
  data = first_diff,
  vcov = 'hc1'
)

print(did_fd)
```

We can verify that the methods all produce the same estimate.

## Question 2: Requirements for Causual Interpretation

Key assumptions needed for causal interpretation of these estimates are:

1.  Parallel Trends Assumption = Treated and control counties would have followed the same trend in the absence of the treatment(tva)

2.  No Spillover Effects = TVA treatment didn't affect manufacutring employment in control counties

3.  No Anticipation Effects = Behavior of counties didn't change before treatment because of anticiaption of TVA

4.  Stable Unit Treatment Assumption = Each country's outcomes only depends on its own treatment status

## Question 3: Pre-trends and 1950 Estimates

```{r pretrends}
calc_did <- function(start_year, end_year) {
  temp_data <- tva |>
    filter(year %in% c(start_year, end_year)) |>
    mutate(
      .by = county_code,
      delta_ln_manufacturing = 
        ln_manufacturing[year == end_year] -
        ln_manufacturing[year == start_year]
    ) |>
    filter(year == end_year)
  
  model <- feols(
    delta_ln_manufacturing ~ i(tva, ref = 0),
    data = temp_data,
    vcov = "HC1"
  )
  
  return(model)
}

did_1920 <- calc_did(1940, 1920) # pre-trend
did_1930 <- calc_did(1940, 1930) # pre-trend
did_1950 <- calc_did(1940, 1950)

print("Pre-trend 1920-1940:")
print(did_1920)
print("Pre-trend 1930-1940:")
print(did_1930)
print("Effect 1940-1950:")
print(did_1950)
```

## Question 4: Event Study

```{r event-study}
tva_event <- tva |>
  mutate(
    event_time = if_else(
      tva == 0,
      -10,  # for untreated group
      year - 1950  # for treated group
    )
  )

event_study <- feols(
  ln_manufacturing ~
    i(tva, i.event_time, ref = 0, ref2 = -10) |
    county_code + year,
  data = tva_event,
  vcov = 'hc1'
)

print(event_study)
```

We can verify the methods from questions 3 and 4 produce the same results.

## Conditional Parallel Trends

## Question 1: Argument for Including Baseline Manufacturing Share

Including baseline manufacturing share as co-variate is important because:

1.  Captures pre-existing industrial development
2.  Areas with different initial manufacturing levels could have different growth movement
3.  Helps control for mean reversion in manufacturing employment
4.  Accounts for potential convergence patterns in regional development

## Question 2: Regression Adjustment Estimator

```{r reg-adjustment}
covariates <- c(
  "agriculture_share_1920", "agriculture_share_1930",
  "manufacturing_share_1920", "manufacturing_share_1930",
  "white_share_1920", "white_share_1930"
)

reg_adj <- feols(
  delta_ln_manufacturing ~ i(tva, ref = 0) + 
    agriculture_share_1920 + agriculture_share_1930 +
    manufacturing_share_1920 + manufacturing_share_1930 +
    white_share_1920 + white_share_1930,
  data = first_diff
)

etable(reg_adj)
```

## Question 3: DRDID Panel Regression

```{r drdid_panel}
drdid_data <- tva_short |>
  arrange(county_code, year)

y1 <- drdid_data$ln_manufacturing[drdid_data$year == 1960]
y0 <- drdid_data$ln_manufacturing[drdid_data$year == 1940]
D <- drdid_data$tva[drdid_data$year == 1960]
covariates <- as.matrix(drdid_data[drdid_data$year == 1960, 
                                   c("agriculture_share_1920", 
                                     "agriculture_share_1930", 
                                     "manufacturing_share_1920", 
                                     "manufacturing_share_1930", 
                                     "white_share_1920", 
                                     "white_share_1930")])

covariates <- cbind(1, covariates)

drdid_panel <- DRDID::reg_did_panel(
  y1 = y1,
  y0 = y0,
  D = D,
  covariates = covariates,
  boot = TRUE,            # Use bootstrap for inference (optional)
  nboot = 999             # Number of bootstrap repetitions
)

summary(drdid_panel)
```

## Question 4: Doubly-robust Estimator

```{r drdid robust}
# Perform Doubly-Robust DiD
drdid_results <- DRDID::drdid_panel(
  y1 = y1,
  y0 = y0,
  D = D,
  covariates = covariates,
  boot = TRUE,    # Bootstrap for inference
  nboot = 999     # Number of bootstrap repetitions
)

summary(drdid_results)
```
